### 2017-07-29 docs/manuscript.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Learning pleiotropic signatures to construct a causal map of the human phenome},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{amsmath}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Learning pleiotropic signatures to construct a causal map of the human
phenome}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{29 July 2017}

\begin{document}
\maketitle

Gibran Hemani\(^1\), Jack Bowden\(^1\), Philip Haycock\(^1\), Jie
Zheng\(^1\), Oliver Davis\(^1\), Peter Flach, Tom Gaunt\(^{1,*}\),
George Davey Smith\(^{1,*}\)

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  MRC Integrative Epidemiology Unit (IEU) at the University of Bristol,
  School of Social and Community Medicine, Bristol, UK
\end{enumerate}

Correspondence to:
\href{mailto:g.hemani@bristol.ac.uk}{\nolinkurl{g.hemani@bristol.ac.uk}}

* Equal contribution

\subsection{Abstract}\label{abstract}

A major application for genome-wide association studies (GWAS) has been
the emerging field of causal inference using Mendelian randomization
(MR), where the causal effect between a pair of traits can be estimated
using only summary level data. MR depends on SNPs exhibiting vertical
pleiotropy, where the SNP influences an outcome phenotype only through
an exposure phenotype. Issues arise when this assumption is violated due
to SNPs exhibiting horizontal pleiotropy, and many methods have been
developed in an attempt to address this. Here we show that the
mechanisms that underlie horizontal pleiotropy are numerous, and that
instrument selection will be increasingly liable to selecting invalid
instruments as GWAS sample sizes continue to grow. We have developed a
mixture of experts machine learning approach (MR-MoE 1.0) that improves
on both power and false discovery rates over unselective use of any
existing methods. Using the approach, we systematically estimated the
causal effects amongst 2407 phenotypes, generating a working draft of
the causal map of the human phenome. Almost 90\% of causal estimates
indicated some level of horizontal pleiotropy. All results are organised
into a publicly available graph database (\url{http://eve.mrbase.org}).

\subsection{Introduction}\label{introduction}

Mendelian randomization (MR) (1,2) exploits genetic pleiotropy to infer
the causal relationships between phenotypes. Suppose that one trait (the
exposure) causally influences another (the outcome). If a SNP influences
the outcome through the exposure then the SNP is exhibiting vertical
pleiotropy. Such a genetic variant is considered to be an instrumental
variable for the if it only influences the outcome through the exposure
(the exclusion restriction). It can be exploited to mimic a randomised
controlled trial, enabling a causal estimate to be made by comparing the
outcome phenotypes between those individuals that have the
exposure-increasing allele against those who do not. Multiple
independent genetic variants for a particular exposure can be used
jointly to improve causal inference, because a) each variant represents
an independent natural experiment, and an overall causal estimate can be
obtained by meta-analysing the single estimates from each instrument;
and b) potential bias arising from the exclusion restriction principal
can be detected or corrected by evaluating the consistency of effects
across instruments (3--8).

Genome-wide association studies (GWAS) have identified genetic
instrumental variables for thousands of phenotypes (9). Recent
developments in MR have enabled knowledge of instrumental variables to
be applied using only summary level data (known as two-sample MR, 2SMR)
(10). Here, in order to infer the causal effect of an exposure on an
outcome all that is required is an estimate of the genetic effects of
the instrumenting SNP on the exposure, and the corresponding estimate of
the effect on the outcome. This has three major advantages. First, GWAS
summary data are non-disclosive and often publicly available. Second,
causal inference can be made between phenotypes even if they have not
been measured in the same samples, limiting possible causal estimates
only by the availability of GWAS summary data for the traits in question
(11). Third, we can view the causal inference problem through the simple
and widely understood prism of a meta-analysis.

Problems with obtaining unbiased causal effects can arise, however, if
the genetic instruments exhibit horizontal pleiotropy (HP), where they
influence the outcome through a pathway other than the exposure (2). The
extent of this phenomenon is not to be understated, and many methods
have been developed that attempt to reliably obtain unbiased causal
estimates under specific models of HP (3--6,8,12). It is considered best
practice to report estimates from all available methods as sensitivity
analyses when presenting causal estimates because different pleiotropic
models can then be evaluated. However it is also desirable to
consolidate across methods under some circumstances. First, it could be
of interest to make causal effect estimates for thousands of traits, in
which case a critical evaluation of each causal estimate of interest may
not be possible or convenient. Second, though the IVW approach is most
statistically powerful under no HP, it can have high false negative or
low true positive rates in the presence of HP compared to other methods.
Pleiotropy has been hypothesised to be universal (13), though the degree
to which this may be the case is contested (14,15), hence defaulting to
the IVW method in the first instance and using other methods as
sensitivity analyses may not be appropriate. Third, if different methods
disagree it is not possible to know which is correct because the true
nature of HP exhibited by the instruments is not known. Fourth, the
available methods do not cover all possible models of HP, and therefore
an automated method for instrument selection may be necessary.

In this paper we introduce two innovations towards improving the
reliability of MR estimates. First, we introduce an approach to discard
genetic variants that are likely to be invalid. Second, we hypothesised
that characteristics of the summary data could indicate which method
would be most reliable, and we introduce new machine learning approaches
that attempt to automate both instrument and method selection. Using
curated GWAS summary data for thousands of phenotypes (11), we use these
new methods to construct a graph of millions of causal estimates.
Motivated by the recent avalanche (16) of 2SMR publications, often with
repeating (17,18) or contradicting (19,20) conclusions despite using the
same data, we developed a graph database to represent these estimates in
a consistent manner. We consider this to be a `working draft' of the
causal map of the human phenome, but raise caution throughout that its
interpretation is far from straightforward, and that future corrections
and refinements will inevitably follow as data grows and 2SMR methods
evolve.

\subsection{Methods}\label{methods}

\subsubsection{GWAS summary data and their use in
2SMR}\label{gwas-summary-data-and-their-use-in-2smr}

The use of summary data in two-sample MR is described in detail
elsewhere (11). A brief outline of the procedure is as follows. First,
genetic instruments for the exposure trait need to be identified - those
SNPs with \(p < 10^{-8}\) are retained in order to ensure that the first
assumption of MR (that the instrument associates with the exposure) is
generally satisfied. We collect their effect sizes, standard errors and
effect alleles for the association with the exposure trait. These can be
obtained manually from complete summary data or from curated lists of
significant GWAS associations. Next, the effects of those SNPs on the
outcome need to be obtained, typically necessitating complete summary
data because these SNPs will generally have not have reached genome-wide
significance for other traits (and therefore be present in curated
catalogues). At its most simple implementation, the regression of the
SNP-exposure effect sizes against the SNP-outcome effect sizes, with
greater weight afforded to those SNPs with smaller SNP-outcome standard
errors, provides the estimate of the causal effect of the exposure on
the outcome. This is known as the inverse-variance weighted (IVW)
method.

Given summary data for a large number of traits, it is straightforward
to exhaustively analyse the causal relationships of every trait against
every other trait for which there are sufficient summary data available.
Supplementary table 1 provides a list of all traits that have available
GWAS summary data, indicating if they have complete summary data (in
which case they can be used as both exposure traits and outcome traits),
or if they only have significant associations (in which they can only be
used as exposure traits).

\subsubsection{MR methods and their
assumptions}\label{mr-methods-and-their-assumptions}

In this paper we consider three main classes of MR estimation. Full
details for each approach have been described previously. A summary of
the methods is given in Supplementary table 2.

\textbf{Mean-based methods:} Here we consider four nested models (5).
The inverse variance weighted (IVW) fixed effects meta-analysis approach
assumes that variants exhibit no HP. IVW random effects meta-analysis
relaxes the HP assumption, allowing it to be present but balanced - such
that it only leads to increased heterogeneity around the regression and
not introducing bias. Fixed effects Egger regression (3) relaxes the HP
assumption further by allowing a non-zero intercept which essentially
allows horizontal pleiotropy to be directional, where it systematically
influences the outcome in a specific direction. Random effects Egger
regression allows heterogeneity around the directional HP (5), as long
as the HP effects are not correlated with the SNP-exposure effects (also
known as the INSIDE assumption)(3).

The Rucker framework (21), adapted to MR (5) uses estimates of
heterogeneity in the IVW and Egger frameworks to navigate between these
nested models. A jackknife approach (random selection with replacement
of the complete set of instruments) can be used to obtain a sampling
distribution for the model estimate amongst these four variations. Using
1000 rounds of jackknife estimates, we can obtain a final estimate using
the mean or the median of the distribution. We only use the jackknife
approach for associations where there are 15 or more variants in order
to avoid saturating the possible number of instrument combinations.

The four nested models (IVW fixed effects, IVW random effects, Egger
fixed effects, Egger random effects) plus the three Rucker estimates
(point estimate, mean of the jackknife, median of the jackknife) provide
seven mean based estimators.

\textbf{Median-based methods:} An alternative approach is to take the
median effect of all available instruments (4,22). This has the
advantage that up to half the instruments can be invalid, and the
estimate will remain unbiased. Developing the approach further to allow
stronger instruments to contribute more towards the estimate can be
obtained by obtaining the median of the weights of each instrument. The
penalised weighted median estimator introduces a further weight to the
instruments, penalising any instrument that contributes substantially
towards the heterogeneity statistic. Together, this provides three
median based estimators. Other estimation strategies not considered
here, such as LASSO regression, have also been developed for the
situation where at least half of the instruments are valid (23).

\textbf{Mode-based methods:} Similar to the median, the mode based
estimator clusters the instruments into groups based on similarity of
causal effects, and returns the final causal effect estimate based on
the cluster that has the largest number of instruments (6). There are
four implementations of this method: the simple and the weighted mode,
each weighted with or without the assumption of no measurement error in
the exposure estimates (NOME). The simple mode is the unweighted mode of
the empirical density function of causal estimates, whereas the weighted
mode is weighted by the inverse variance of the outcome effect. The
bandwidth parameter was set to 1 by default.

\subsubsection{Instrument selection}\label{instrument-selection}

\paragraph{Top hits}\label{top-hits}

The simplest approach to selecting instruments for performing MR is to
take SNPs that have been declared significant in the published GWAS for
the exposure. This typically involves obtaining SNPs that surpass
\(p < 5 \times 10^{-8}\), using clumping to obtain independent SNPs, and
then replicating in an independent sample. These results are often
recorded in public GWAS catalogs. Alternatively the clumping procedure
can be performed using complete summary data in MR-Base (11). We call
this the ``top hits'' strategy.

\paragraph{Steiger filtering}\label{steiger-filtering}

With genome-wide association studies growing ever larger, the
statistical power to detect significant associations that may be
influencing the trait downstream of many other pathways increases. For
example, if a SNP \(g_{A}\) influences trait \(A\), and trait \(A\)
influences trait \(B\), then a sufficiently powered GWAS will identify
the \(g_{A}\) as being significant for trait \(B\) (Figure 1a). Using
\(g_{A}\) as an instrument to test the causal effect of \(A\) on \(B\)
is perfectly valid. But in the (incorrectly hypothesised) MR analysis of
trait \(B\) on trait \(A\) could erroneously result in the apparent
causal association of \(B\) on \(A\). If \(g_{A}\) is only one of many
known instruments for \(B\), amongst which some are valid, it is to the
advantage of the researcher to exclude \(g_{A}\) from the analysis.

An approach to inferring the causal direction between phenotypes
recently developed (7) uses the following basic premise. If trait \(A\)
causes trait \(B\) then

\[
\sum^M_{i=1}{cor(g_{i}, A)^2} > \sum^M_{i=1}{cor(g_{i}, B)^2}
\]

because the \(cor(g_{i}, B)^2 = cor(A, B)^{2} cor(g_{i}, A)^{2}\). This
simple inequality will not hold in some cases, for example
\(\rho_{x, x_o} < \rho_{x,y}\rho_{y,y_o}\) where \(\rho_{x, x_o}\) and
\(\rho_{y, y_o}\) are the precision of the measurements of the \(x\) and
\(y\). Some combinations of confounding effects can also distort the
\(\rho_{g,x}\) and \(\rho_{g,y}\) parameters, as has been discussed in
detail previously. However, we use it here as a computationally
inexpensive and approximate method to identify variants that are likely
to be invalid. Steiger's Z-test of correlated correlations (24) can be
used to formally test the extent to which the two correlations are
statistically different.

Here we adapt this approach to automatically filter SNPs that are liable
to be invalid (Figure 1a). Other methods have been developed to identify
invalid instruments for the purposes of exclusion from MR analyses
(8,12,25), based on the notion that outlying instruments are likely to
be exhibiting horizontal pleiotropy. Conversely, we primarily developed
Steiger filtering to identify those instruments that are likely to be
arising due to reverse cause.

The Steiger test is applied to each variant in turn and we exclude any
\(g_{A}\) for which \(cor(g_{i}, A)^2 > cor(g_{i}, B)^2\), indicating
that it is unlikely to primarily associate with \(B\) relative to \(A\).
Similarly, for SNPs that influence confounders of \(A\) and \(B\) or
exhibit horizontal pleiotropy, the difference in \(cor(g_{i}, A)^2\) and
\(cor(g_{i}, B)^2\) will be reduced, increasing the likelihood of the
SNP being excluded because the Steiger Z-test is less likely to be
significant. Hence we also exclude any \(g_{A}\) for which the Steiger
Z-test has \(p > 0.05\).

To estimate \(cor(g, x)^2\), if \(x\) is continuous we obtain the
F-statistic from the reported p-value and sample size and then
\(cor(g, x)^2 = \frac{F}{N - 2 - F}\). If \(x\) is binary then we
estimate the variance of the underlying liability explained by the SNP,
\(cor(g, x)^2 = \frac{V_a}{V_a + V_e}\). Here, \(V_e = \pi^2/3\), and
\(V_a = 2\beta^2p(1-p)\), where \(\beta\) is the log odds ratio and
\(p\) is the allele frequency of the SNP in the population (26). \(p\)
can be estimated using the allele frequency of the SNP in an ascertained
sample by deriving the \(2 \times 2\) contingency table from the odds
ratio \(e^\beta\), allele frequency in the ascertained sample
\$p\_\{cc\}, and number of cases \(N_1\) and controls \(N_0\). We
exclude any SNPs that do not have the correct causal direction for the
hypothesis, and do not have Steiger test \(p < 0.05\).

\subsubsection{Competitive mixture of
experts}\label{competitive-mixture-of-experts}

We consider the 14 MR methods described above, for which instruments can
be supplied using two instrument selection strategies, leading to 28
methods in total (Supplementary table 2). In the context of this
analysis, each method is considered to be an `expert', taking a set of
SNP-exposure and SNP-outcome effect sizes and their standard errors as
inputs. Our objective is to select the expert most likely to be correct
for a specific MR analysis. A overview of the approach is shown in
Figure 2.

The mixture of experts method is a machine learning approach (27) which
seeks to divide a parameter space into subdomains, such that a
particular expert is used primarily for problems that reside in a
subdomain most suited to that expert. In this case our objective is to
identify characteristics of the SNP-exposure and SNP-outcome
associations for which one specific MR method is most likely to yield
highest statistical power for non-null associations, and lowest false
discovery rates for null associations. This involves creating a `gating
function', whose purpose is to decide which expert to use for a specific
MR analysis, given the parameter space that is occupied by that dataset.
The metrics are a mixture of regression diagnostics and are described in
Supplementary table 3.

The gating function needs to be trained using data for which the true
causal effect is known, and to this end we generated a large number of
simulated datasets. We trained the gating function using random forest
learning algorithms that seek to identify sectors of the parameter space
that are most likely to return accurate estimates for a particular
expert. Figure 2a illustrates how the gating function is trained using
simulated data. New datasets can then be applied to the trained mixture
of experts (MoE) model (Figure 2b). We call this implementation MR-MoE
1.0.

It should be noted that in the original hierarchical mixture of experts
approach of (27) the gating function and the experts share the same
input space and are trained simultaneously using
expectation-maximisation. In our approach the gating function is defined
over a separate input space consisting of the parameter space of the
experts, and is trained separately in a supervised setting using
simulated data. As such our MR-MoE approach fits the framework of
meta-learning (28--30) where machine learning methods are applied at the
`meta-level' to analyse the results of machine learning experiments at
the `base-level', in order to be able to understand and model the
capabilities of each expert and recommend which expert should be applied
to a given problem. From this perspective the 28 MR methods are the
base-models, the gating function is the meta-model, and its
meta-features are the parameters of the datasets.

\textbf{Training and testing simulations}

The MoE is trained using datasets generated from simulations (Figure
2a). A \emph{dataset} is the minimum data required to perform 2SMR
analysis - four columns comprising the SNP-exposure and SNP-outcome
effects and standard errors, and rows corresponding to each SNP that is
used as an instrument for the exposure. This \emph{dataset} can be fed
into any of the 28 experts to obtain MR causal effects. The simulations
used to generate these datasets seek to cover a range of pleiotropic
scenarios, including where some proportion of SNPs exhibit directional
or balanced horizontal pleiotropy, or where SNPs influence confounding
variables.

We simulate two individual level datasets for which there are \(N_x\)
and \(N_y\) samples, and \(M\) SNPs, where each SNP has effect allele
frequency of \(p_m \sim U(0.05, 0.95)\). These datasets are used to
obtain the SNP effects for the exposure trait \(x\) and the outcome
trait \(y\), respectively, using the following sampling criteria:

\[
\begin{aligned}
N_x & = \{20000, ..., 500000\} \\
N_y & = \{20000, ..., 500000\} \\
K & = \{0, ..., 10\} \\
M_x & = \{1, ..., 200\} \\
M_y & = \{1, ..., 200\} \\
M_{u_k} & = \{5,..., 30\} \\
\end{aligned}
\]

The \(M = M_x + M_y + \sum{M_{u_k}}\) SNPs can influence \(x\) directly,
\(y\) directly, or some number of confounders \(u_{k}\) directly.
Phenotypes for \(x\) and \(y\) are constructed using

\[
x = \sum^{M_x}_{i}{\beta_{gx,x,i}g_{x,i}} + \sum^{M_y}_{j}{\beta_{gy,x,j}g_{y,j}} + \sum^{K}_{k}{\beta_{ux,k} u_{k}} + e_{x}
\]

where \(\beta_{gx,x}\) is the vector of effects of each of the \(M_x\)
SNPs that influence \(x\) primarily, \(\beta_{gy,x}\) is the vector of
effects for the \(M_y\) SNPs on \(x\), where the \(M_y\) SNPs influence
\(y\) primarily but exhibit horizontal pleiotropic effects on \(x\). We
allow some proportion of these effects to be 0. \(\beta_{ux}\) is the
vector of effects of each of the \(K\) confounders on \(x\). Each
\(u_{k}\) variable is constructed using

\[
u = \sum^{M_u}_{l}{\beta_{gu,l}g_{l}} + e_{l}
\]

and finally \(y\) is constructed using

\[
y = \beta_{x,y}x + \sum^{M_y}_{i}{\beta_{gy,y,j}g_{y,j}} + \sum^{M_x}_{j}{\beta_{gx,y,i}g_{x,i}} + \sum^{K}_{k}{\beta_{uy,k} u_{k}} + e_{y}
\]

where \(\beta_{x,y}\) is the causal effect of \(x\) on \(y\). We sample
the distribution of direct SNP effects using

\[
\begin{aligned}
\beta_{gx,x,i} & \sim N(0, \sigma^2_{gx,x}) \\
\sigma^2_{gx,x,i} & \sim U(0.01, 0.1) \\
\beta_{gy,y,j} ~ N(0, \sigma^2_{gy,y}) \\
\sigma^2_{gy,y,j} & \sim U(0.01, 0.1) \\
\end{aligned}
\]

Some proportion \(floor(MS_{x}/M)\) of \(g_x\) SNPs and
\(floor(MS_{y}/M)\) of \(g_y\) SNPs, where \(s_x\) and
\(s_y \sim U(0,1)\), exhibit horizontal pleiotropy with effects sampled
using

\[
\begin{aligned}
\beta_{gx,y,i*} & \sim N(\mu_{gx,y}, \sigma^2_{gx,y})  \\
\mu_{gx,y,i*} & \sim U(-0.005, 0.005) \\
\sigma^2_{gx,y,i*} & \sim U(0.001, 0.01) \\
\beta_{gy,x,j*} & \sim N(\mu_{gy,x}, \sigma^2_{gy,x}) \\
\mu_{gy,x,j*} & \sim U(-0.005, 0.005) \\
\sigma^2_{gy,x,j*} & \sim U(0.001, 0.01) \\
\end{aligned}
\]

The genetic influences on each of the confounders are sampled using

\[
\begin{aligned}
\beta_{gu,u,l} & \sim N(0, \sigma^2_{gu,u}) \\
\sigma^2_{gu,u,l} & \sim U(0.01, 0.1) \\
\end{aligned}
\]

The influence of each confounder on \(x\) and \(y\) is obtained using

\[
\begin{aligned}
\beta_{u,x} & \sim N(0, \sigma^{2}_{u,x}) \\
\beta_{u,y} & \sim N(0, \sigma^{2}_{u,y}) \\
\end{aligned}
\]

Finally, 20\% of the simulations have a null effect of
\(\beta_{x,y} = 0\), while the other remaining 80\% have a true effect
sampled from

\[
\begin{aligned}.
\beta_{x,y} & \sim N(0, \sigma^2_{x,y}) \\
\sigma^2_{x,y} & \sim U(0.001, 0.1) \\
\end{aligned}
\]

For each simulation we used linear regression to estimate the genetic
effect of each SNP \(M\) on \(x\) in sample 1, and each SNP \(M\) on
\(y\) in sample 2. We then perform MR analysis in both directions,
mimicking GWAS by retaining SNPs that have \(p < 5e-8\) in sample 1 to
perform MR of \(x\) on \(y\) (the true causal direction for non-null
simulations), and retaining SNPs that have \(p < 5e-8\) in sample 2 to
perform MR of \(y\) on \(x\) (the reverse causal direction for non-null
simulations). We treat the summary data (effect sizes and standard
errors) used for estimating \(x \rightarrow y\) the summary data used
for estimating \(y \rightarrow x\) as two separate datasets Hence, for
each simulation two datasets are generated which are analysed to produce
28 MR estimates each. We performed 100,000 simulations using these
parameters, resulting in 200,000 datasets.

\textbf{Optimisation function}

We aim to maximise statistical power for datasets where
\(\beta_{x,y} \neq 0\) and minimise false discovery rates for datasets
where \(\beta_{x,y} = 0\). To train random forest decision trees to
predict performance for a particular method
\(h(O_{w,d}, \textbf{z}_{d})\) is generated where the training set of
input metrics for dataset \(d\) is \(\textbf{z}_{d}\) and the response
(optimisation function) is

\[
    O_{w,d} = 
\begin{cases}
    1,   & \text{if } \beta_{x,y} \neq 0 \text{ and } p_{m,d} < 0.01\\
    1,   & \text{if } \beta_{x,y} = 0 \text{ and } p_{m,d} > 0.1 \\
    0,   & \text{otherwise}
\end{cases}
\]

where \(p_{m,d}\) is the p-value for method \(m\) on dataset \(d\).

\textbf{Strategy}

For each training dataset we record a set of 53 metrics
\(\textbf{z}_{d}\) (Supplementary table 3), and an outcome \(O_{w,d}\),
which is a measure of how well that method performed for each particular
simulated dataset. For each of our 28 methods, we need to create a model
that predicts the performance of the method based on metrics generated
from a dataset. To do this, for each method we train a random decision
forest to predict that method's performance using the dataset's metrics.
The random forest approach is well suited to this problem because there
are likely to be non-continuous combinations of different metrics that
improve on prediction over, for example, a simple linear model that does
not learn about interactions.

As a simple hypothetical example - if a dataset has a single outlier but
is otherwise exhibiting no heterogeneity then the following methods
could arguably perform well:

\begin{itemize}
\tightlist
\item
  an IVW fixed effects analysis with the outlier removed, should the
  Steiger test be able to detect the outlier
\item
  a median based approach
\item
  a mode based approach
\end{itemize}

deciding between these methods requires finding, in general, which will
minimise false discovery rates and maximise true positive rates for that
particular scenario. In this example the IVW with Steiger filtering
would likely be the clear winner because of its superior statistical
power. Countless other scenarios could arise. For example if 100\% of
instruments are invalid but the InSIDE assumptio is met then the
MR-Egger is likely to be most effective; if 40\% of instruments are
invalid then a median-based approach is likely most effective; and if
80\% of instruments are invalid then a mode-based approach is likely
most effective.

Having generated random forest decision trees for each of the 28 methods
using 133,000 of the simulations, we then applied them to the remaining
67,000 datasets to predict which method would have the highest
performance for each of the remaining datasets. Finally we compare the
performance of the method selected by the MoE against all remaining
methods. The default settings for the \texttt{randomForest} package in R
(31) were used to train the models. MR-MoE 1.0 is implemented in the
TwoSampleMR R package available at
\href{https://github.com/MRCIEU/TwoSampleMR}{github.com/MRCIEU/TwoSampleMR}
(11).

\subsubsection{Graph database of MR
estimates}\label{graph-database-of-mr-estimates}

The set of MR estimates obtained from this analysis are recorded in a
Neo4j graph database. Because each association has up to 28 different
estimates, for simplicity we distill this down to a single `best
estimate' for each association using the following rules:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  If the number of variants after Steiger filtering is greater than 5
  then apply the MoE to obtain the best method
\item
  If the number of variants after Steiger filtering is less than or
  equal to 5 but greater than 1 then use the IVW random effects approach
  on the filtered set of variants
\item
  If there is 1 variant retained in the Steiger approach then use the
  Wald ratio on the remaining variant
\item
  If there are no variants remaining after Steiger filtering then
  declare no causal association.
\end{enumerate}

For specific hypotheses we strongly recommend that estimates from all
sensitivity analyses are scrutinised and reported. The graph can be
queried directly using the cypher language at
\url{http://eve-neo4j.mrbase.org} or through a basic web interface at
\url{http://eve.mrbase.org}.

\subsection{Results}\label{results}

\subsubsection{Steiger filtering improves
reliability}\label{steiger-filtering-improves-reliability}

As statistical power for GWAS studies improves the likelihood of a
significant association being discovered that doesn't act primarily on
the trait of interest increases (Figure 1a). We evaluated the efficacy
of the Steiger filtering approach for improving instrument selection
using 100,000 simulated datasets comprising both null and non-null
causal models. For each dataset, instruments were selected based on the
tophits strategy and the Steiger filtering strategy, and MR was
performed using 14 different methods based on instruments selected from
each of these strategies.

Figure 1b shows that the tophits strategy led to over half of the
instruments being primarily associated with either confounders or the
outcome phenotype, not the exposure phenotype. The proportion of invalid
instruments due to reverse cause increased as GWAS discovery sample size
increased. Applying Steiger filtering reduced this to 25\%.
Consequently, the false discovery rates (FDR) for 12 of the 14 methods
reduced substantially when applied using Steiger filtered instruments
(Figure 1c). The true positive rates for the methods based on Steiger
filtering did however reduce slightly for 10 of the 14 methods.

\subsubsection{Mixture of experts method selection improves over any
single
method}\label{mixture-of-experts-method-selection-improves-over-any-single-method}

Following evidence that the Steiger filtering approach can improve on
existing methods, we next hypothesised that a mixture of experts (MoE)
model would be able to predict the most appropriate of the 28 MR methods
and instrument selection strategies to apply to a particular dataset
based on its characteristics (Figure 2).

The ability to predict the performance for each of these methods is
shown in Supplementary table 2. The prediction \(R^2\) of whether a
dataset's status was truly null (\(\beta_{xy}=0\)) or non-null
(\(\beta_{xy} \neq 0\)) against the method's prediction of the dataset's
status, ranged between 0.04 and 0.24. The method performance prediction
was most effective for the Egger random effects model with Steiger
filtering. The dataset characteristics with the most importance for each
of the predictors differed substantially between each dataset, as well
as the frequency for which each of the methods was selected in the
testing datasets. The FDR of each method when chosen, compared to their
averages across all datasets, reduced; and likewise the true positive
rates of each method when chosen increased compared to their
simulation-wide averages.

We compared the MoE performance in the simulations against each of the
28 methods, testing to see if it outperformed all other single methods.
Figures 3a and 3b show that the MoE approach had the best general
performance, obtaining an AUC of 0.84 in terms of classifying the
simulations as being null or non-null. Notably, the next best methods
were median and mode based estimators using Steiger filtering. Under the
assumption of widespread pleiotropy it is clear that all methods suffer
from high false discovery rates.

\subsubsection{Automated MR analysis of 2407
phenotypes}\label{automated-mr-analysis-of-2407-phenotypes}

We applied our analysis using summary data for 2407 phenotypes,
including 149 complex traits and diseases, 575 metabolites (32,33) and
1683 plasma protein levels (34). For the protein levels only the
instruments were available, so they could only be evaluated as exposure
phenotypes. The complex traits and diseases and metabolite levels had
complete summary data available in MR-Base, thus they be evaluated as
both exposures (if they had significant instruments) and outcomes.
Together, we evaluated 715681 relationships. The majority of these
associations could only be evaluated using fewer than 5 SNPs, and so the
Wald ratio or IVW fixed effects methods were used, but for 61029
associations the MR-MoE approach was applied.

There were 5660 associations following Bonferroni correction
(\(p < 7.0 \times 10^{-8}\)). Of these 2918 were obtained from the
MR-MoE analysis, while the remainder were estimated using fewer than 5
SNPs using the Wald ratio or IVW fixed effects methods.

The frequencies of the methods chosen by the MR-MoE analysis are shown
in Supplementary table 3. Amongst those deemed `significant', the IVW
fixed effects analysis using tophits instruments (the only method
applicable when there is no evidence of horizontal pleiotropy of any
sort) was selected in only 10.4\% of cases. This indicates that
horizontal pleiotropy is likely to be pervasive.

\subsubsection{Putative associations involving years of
schooling}\label{putative-associations-involving-years-of-schooling}

As an example of the database we performed a look up of associations
where LDL cholesterol as measured in the GLGC consortium (35) had a
false discovery rate of 0.05 using the MoE approach. This returned 287
putative associations, amongst which 111 involved LDL cholesterol
influencing other traits and 176 involved other traits influencing LDL
cholesterol. A large proportion of these traits were metabolites, which
are dominated by lipid fractions. Filtering to exclude the metabolomic
studies ((32,33)) returned 27 associations with 23 traits (Figure 4a).
There was a strong association with coronary heart disease (0.41 log(OR)
per SD, 0.31-0.52, \(p=2.7\times10^{-12}\)). Here the IVW random effects
method was chosen after Steiger filtering, indicating a prediction of HP
being present amongst the instruments.

We also performed a look up of associations that causally influence
years of schooling (36), or that years of schooling itself causally
influences. Using a false-discovery rate of 0.05, 45 traits were
returned as having some direction of causality with years of schooling
(Figure 4b). Several of these putative relationships appear to be
plausible, for example the influence of years of schooling on lower risk
of Alzheimer's disease has recently been examined (37). However it is
clear that there are serious limitations in relying on MR as a panacea
for causal inference. For example, bi-directional causal relationships
with college completion exist because of the similarity in trait
definitions; and there are several traits (e.g.~childhood intelligence
and birth weight) which appear to be causally downstream of years of
schooling despite this being temporally impossible. Though explanations
could be conjured for this association, for example parents' schooling
influences childhood intelligence, or the genetic instruments are shared
across the two traits, an intuitive interpretation of causality cannot
be applied.

The facility to perform these scans has been automated in the MR of
everything vs everything (MR-EvE) web application,
\url{http://eve.mrbase.org}.

\subsection{Discussion}\label{discussion}

The trend of increasing GWAS sample sizes continues, but while the
opportunity that this affords MR to be furnished with more instruments
is typically welcomed, here we have shown that it is invalid
instruments, exhibiting horizontal pleiotropy or reverse causation, that
are more likely to be identified than valid ones. Strategising how MR is
to be used in practice, therefore, must consider horizontal pleiotropy
to be the rule rather than the exception. Following a recent flurry in
development of MR methodology, we have devised a machine learning
approach, MR-MoE, that seeks to predict the performance of each MR
method, selecting the one which is most likely to maximise power whilst
minimising FDR for a specific dataset. The mixture of experts approach
that we present here is trained using random forest decision trees
applied to extensive and diverse simulations, and we demonstrate that it
makes substantial performance improvements over any other single method
in the presence or absence of extensive horizontal pleiotropy. As such,
our method contributes to the field of meta-learning but also to the
rapidly growing field of automatic model selection and hyper-parameter
optimization in machine learning (38).

Crucially however, we observed in our simulations that none of the
available methods were conservative and all, including MR-MoE, had FDR
above 0.2. Instrumental variables in MR are typically chosen blind, with
GWAS significance being the only criteria. We illustrated how
confounders and reverse causal associations can easily lead to invalid
instruments being used in MR, and that even if those instruments that
only influence the exposure directly are used, most randomly generated
patterns of horizontal pleiotropy cannot be adequately accounted for by
any single MR method.

We applied the method to the curated set of GWAS summary data present in
MR-Base (11), generating several million MR estimates. MR-MoE selected a
method that indicated some pattern of horizontal pleiotropy for almost
90\% of cases, reinforcing the notion that it is the rule rather than
the exception. Based on our simulations we expect that those
associations that we reported to be `significant' are liable to high
type 1 error rates.

In addition to correctly handling horizontal pleiotropy, there are many
other limitations that prevent MR from being a panacea for causal
inference. Many of the associations are biologically impossible, for
example where early life phenotypes appear to be influenced by later
stage phenotypes. Though these associations can be informative, their
interpretations as causal relationships are far from clear. Similarly,
often disease traits appear to causally relate to other phenotypes, but
GWAS is typically performed on the liability scale, hence the causal
estimate reflects not the presence or absence of disease, but the
underlying risk of disease. Again, interpretation of such associations
can be problematic.

A large proportion of the associations that were estimated used only a
single instrumental variable. Methods are emerging to attempt to
delineate between models of reverse cause, pleiotropy and multiple
causal variants in the same region (39,40). Separating vertical from
horizontal pleiotropy with a single instrument is not yet possible in
the two-sample MR framework though the use of mediation-based approaches
may be informative in the one-sample setting (7,41). Other problems can
also manifest, for example frailty effects could induce associations for
late-onset traits (42); genetic variants could strongly relate to
several phenotypes making it difficult to ascertain which of them is the
causal exposure (43); and the measured feature for which genetic
associations are known could relate in complicated ways to the
biological entity that are truly causal (2).

We also encounter a potential new problem in using machine learning
approaches to infer the correct model for a particular dataset, namely,
that the data is choosing the model. This is liable to contribute to
elevated type 1 error rates, even though we have attempted to separate
the information used for optimisation from the information used to
predict performance. The extent to which a machine based method selector
suffers from this more than a human manually inferring the result most
likely to be true is not clear. Though MR-MoE does exhibit higher type 1
error rates than are desirable, they still remain amongst the lowest
compared to all other MR methods that do not suffer from this issue.

The field of MR is evolving. The advent of GWAS databases (9,44) and
automated 2SMR (11) has trivialised the analytical aspect of
investigating specific causal hypotheses. Despite the limitations
described above, the construction of a causal graph of `everything
versus everything' does have appeal. First, though causality is not
guaranteed by MR, it can still be highly informative for confirming or
negating specific hypotheses. Second, it paves the way for new
approaches to search for novel putative associations and improve
reliability in single analyses. The potential to exploit GWAS summary
data within the properties of graph databases to aid with both of these
endeavours could be transformative in biological research.

\newpage

\includegraphics{images/fig1.pdf}

Figure 1: Simulations. a) Schematic of how GWAS with sufficient power
can lead to the selection of instrumental variables that are invalid. We
used arbitrary numbers of SNPs and confounders to simulate GWAS summary
datasets. b) Any SNP that has a direct influence on the exposure, or an
influence on a non-confounding intermediate variable, is considered a
`direct' effect. The y-axis shows the proportion of instruments selected
for analysis due that are either direct associations with the exposure,
instruments for the outcome (reverse), or instruments for confounding
traits (confounder). The proportions are compared over a range of
different exposure discovery sample sizes (x-axis) and using either the
tophits approach (left) or the Steiger approach (right) for instrument
selection. c) Top: The false discovery rates from null simulations for
each of the 14 methods using either tophits, Steiger filtered variants,
or variants that are known to be directly associated with the exposure
(oracle, note that direct effects can still exhibit horizontal
pleiotropy in these simulations). Bottom: The statistical power to
detect true causal associations in the non-null simulations.

\newpage

\includegraphics{images/fig2.pdf}

Figure 2: Mixture of experts. a) Training. Datasets are simulated that
have either null or non-null causal relationships, and MR is performed
by each of the 28 available MR methods. In the toy datasets, the columns
in black represent 53 metrics about each of the 67,000 training
simulations. The columns in red are specific to each method, they
represent how well that method performed in obtaining the correct answer
for each of the datasets. Random forests are used to learn the parameter
space of the 53 metrics in which a particular dataset is likely to
perform well. Together, this creates 28 random forest decision trees,
one for each method. b) Application. For a GWAS summary dataset, our
objective is to choose the method most likely to return the correct
causal estimate. Metrics are generated from the dataset and fed into
each of the 28 random forest decision trees. This provides us with 28
performance predictions. Finally, we use the method for which the
performance prediction is highest.

\newpage

\includegraphics{images/fig3.pdf}

Figure 3: Performance of MoE against all other methods. a) The power for
non-null datasets is plotted against the FDR for null datasets for each
of the 28 methods, plus MoE. No single method achieved nominal FDR for
these simulations. b) Calculating the area under the ROC curve from the
values in (a) we plotted the performance in order from lowest to
highest. Under the assumption of pervasive horizontal pleiotropy, the
MoE approach is likely most effective than any other single method.

\newpage

\includegraphics{images/fig4-01.png}

Figure 4: Lookup of causal associations involving a) LDL cholesterol and
b) Years of Schooling. In a) the nodes are filtered to not include those
obtained from the Shin et al or Kettunen et al metabolomic studies. The
arrows denote causal direction and the values on the arrows denote the
causal effect estimate. Only those relationships are shown for which the
\(FDR < 0.05\).

\newpage

\subsection{References}\label{references}

\raggedright

\hypertarget{refs}{}
\hypertarget{ref-DaveySmith2003}{}
1. Davey Smith G, Ebrahim S. 'Mendelian randomization': can genetic
epidemiology contribute to understanding environmental determinants of
disease? International Journal of Epidemiology {[}Internet{]}. 2003
Feb;32(1):1--22. Available from:
\url{http://www.ije.oxfordjournals.org/cgi/doi/10.1093/ije/dyg070}

\hypertarget{ref-DaveySmithHemani2014}{}
2. Davey Smith G, Hemani G. Mendelian randomization: genetic anchors for
causal inference in epidemiological studies. Human molecular genetics.
2014 Jul;23(R1):R89-----R98.

\hypertarget{ref-Bowden2015}{}
3. Bowden J, Davey Smith G, Burgess S. Mendelian randomization with
invalid instruments: effect estimation and bias detection through Egger
regression. International Journal of Epidemiology. 2015;44(2):512--25.

\hypertarget{ref-Bowden2016b}{}
4. Bowden J, Davey Smith G, Haycock PC, Burgess S. Consistent Estimation
in Mendelian Randomization with Some Invalid Instruments Using a
Weighted Median Estimator. Genetic Epidemiology {[}Internet{]}. 2016
May;40(4):304--14. Available from:
\href{http://www.ncbi.nlm.nih.gov/pubmed/27061298\%20http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC4849733\%20http://doi.wiley.com/10.1002/gepi.21965}{http://www.ncbi.nlm.nih.gov/pubmed/27061298 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC4849733 http://doi.wiley.com/10.1002/gepi.21965}

\hypertarget{ref-Bowden2017}{}
5. Bowden J, Del~Greco~M F, Minelli C, Davey Smith G, Sheehan N,
Thompson J. A framework for the investigation of pleiotropy in
two-sample summary data Mendelian randomization. Statistics in Medicine
{[}Internet{]}. 2017; Available from:
\url{http://doi.wiley.com/10.1002/sim.7221}

\hypertarget{ref-Hartwig2017}{}
6. Hartwig FP, Davey Smith G, Bowden J. Robust Inference In Two-Sample
Mendelian Randomisation Via The Zero Modal Pleiotropy Assumption.
bioRxiv {[}Internet{]}. 2017; Available from:
\url{http://biorxiv.org/content/early/2017/04/10/126102}

\hypertarget{ref-Hemani2017}{}
7. Hemani G, Tilling K, Davey Smith G. Orienting The Causal Relationship
Between Imprecisely Measured Traits Using Genetic Instruments. bioRxiv
{[}Internet{]}. 2017; Available from:
\url{http://biorxiv.org/content/early/2017/03/15/117101}

\hypertarget{ref-Verbanck2017}{}
8. Verbanck M, Chen C-Y, Neale B, Do R. Widespread pleiotropy confounds
causal relationships between complex traits and diseases inferred from
Mendelian randomization. bioRxiv {[}Internet{]}. 2017; Available from:
\url{http://www.biorxiv.org/content/early/2017/06/30/157552}

\hypertarget{ref-Hindorff2010}{}
9. Hindorff LA, Junkins HA, Hall PN, Mehta JP, Manolio TA. A Catalog of
Published Genome-Wide Association Studies, available at
http://www.genome.gov/gwastudies. Accessed 12/10/2010. 2010.

\hypertarget{ref-Pierce2013}{}
10. Pierce BL, Burgess S. Efficient design for Mendelian randomization
studies: subsample and 2-sample instrumental variable estimators.
American journal of epidemiology {[}Internet{]}. 2013
Oct;178(7):1177--84. Available from:
\href{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3783091\%7B/\&\%7Dtool=pmcentrez\%7B/\&\%7Drendertype=abstract}{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3783091\{\textbackslash{}\&\}tool=pmcentrez\{\textbackslash{}\&\}rendertype=abstract}

\hypertarget{ref-Hemani2016}{}
11. Hemani G, Zheng J, Wade KH, Laurin C, Elsworth B, Burgess S, et al.
MR-Base: a platform for systematic causal inference across the phenome
using billions of genetic associations. BioRxiv. 2016;10.1101/07.

\hypertarget{ref-Zhu2017}{}
12. Zhu Z, Zheng Z, Zhang F, Wu Y, Trzaskowski M, Maier R, et al. Causal
associations between risk factors and common diseases inferred from GWAS
summary data. bioRxiv {[}Internet{]}. 2017; Available from:
\url{http://www.biorxiv.org/content/early/2017/07/26/168674}

\hypertarget{ref-Wright1968}{}
13. Wright S. Evolution and the Genetics of Populations. In: Evolution
and the genetics of populations. Chicago: University of Chicago Press;
1968.

\hypertarget{ref-Wagner2011}{}
14. Wagner GP, Zhang J. The pleiotropic structure of the
genotype--phenotype map: the evolvability of complex organisms. Nature
Reviews Genetics {[}Internet{]}. 2011 Mar;12(3):204--13. Available from:
\url{http://www.nature.com/doifinder/10.1038/nrg2949}

\hypertarget{ref-Hill2012a}{}
15. Hill WG, Zhang X-S. Assessing pleiotropy and its evolutionary
consequences: pleiotropy is not necessarily limited, nor need it hinder
the evolution of complexity. Nature Reviews Genetics {[}Internet{]}.
2012 Feb;13(4):296. Available from:
\url{http://www.ncbi.nlm.nih.gov/pubmed/22349131}

\hypertarget{ref-Hartwig2016}{}
16. Hartwig FP, Davies NM, Hemani G, Davey Smith G. Two-sample Mendelian
randomization: avoiding the downsides of a powerful, widely applicable
but potentially fallible technique. International Journal of
Epidemiology {[}Internet{]}. 2016 Dec;45(6):1717--26. Available from:
\url{https://academic.oup.com/ije/article-lookup/doi/10.1093/ije/dyx028}

\hypertarget{ref-Xu2017}{}
17. Xu L, Lin SL, Schooling CM. A Mendelian randomization study of the
effect of calcium on coronary artery disease, myocardial infarction and
their risk factors. Scientific Reports {[}Internet{]}. 2017 Feb;7:42691.
Available from:
\href{http://www.ncbi.nlm.nih.gov/pubmed/28195141\%20http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC5307362\%20http://www.nature.com/articles/srep42691}{http://www.ncbi.nlm.nih.gov/pubmed/28195141 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC5307362 http://www.nature.com/articles/srep42691}

\hypertarget{ref-Larsson2017}{}
18. Larsson SC, Burgess S, Michaëlsson K, TB H, WH C, Y P. Association
of Genetic Variants Related to Serum Calcium Levels With Coronary Artery
Disease and Myocardial Infarction. JAMA {[}Internet{]}. 2017
Jul;318(4):371. Available from:
\url{http://jama.jamanetwork.com/article.aspx?doi=10.1001/jama.2017.8981}

\hypertarget{ref-Prins2016}{}
19. Prins BP, Abbasi A, Wong A, Vaez A, Nolte I, Franceschini N, et al.
Investigating the Causal Relationship of C-Reactive Protein with 32
Complex Somatic and Psychiatric Outcomes: A Large-Scale Cross-Consortium
Mendelian Randomization Study. Hay PJ, editor. PLOS Medicine
{[}Internet{]}. 2016 Jun;13(6):e1001976. Available from:
\href{http://www.ncbi.nlm.nih.gov/pubmed/27327646\%20http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC4915710\%20http://dx.plos.org/10.1371/journal.pmed.1001976}{http://www.ncbi.nlm.nih.gov/pubmed/27327646 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC4915710 http://dx.plos.org/10.1371/journal.pmed.1001976}

\hypertarget{ref-Inoshita2016}{}
20. Inoshita M, Numata S, Tajima A, Kinoshita M, Umehara H, Nakataki M,
et al. A significant causal association between C-reactive protein
levels and schizophrenia. Scientific Reports {[}Internet{]}. 2016
Sep;6(1):26105. Available from:
\href{http://www.ncbi.nlm.nih.gov/pubmed/27193331\%20http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC4872134\%20http://www.nature.com/articles/srep26105}{http://www.ncbi.nlm.nih.gov/pubmed/27193331 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC4872134 http://www.nature.com/articles/srep26105}

\hypertarget{ref-Rucker2011}{}
21. Rucker G, Schwarzer G, Carpenter JR, Binder H, Schumacher M.
Treatment-effect estimates adjusted for small-study effects via a limit
meta-analysis. Biostatistics {[}Internet{]}. 2011 Jan;12(1):122--42.
Available from:
\href{http://www.ncbi.nlm.nih.gov/pubmed/20656692\%20https://academic.oup.com/biostatistics/article-lookup/doi/10.1093/biostatistics/kxq046}{http://www.ncbi.nlm.nih.gov/pubmed/20656692 https://academic.oup.com/biostatistics/article-lookup/doi/10.1093/biostatistics/kxq046}

\hypertarget{ref-Han2008}{}
22. Han C. Detecting invalid instruments using L 1-GMM. Economics
Letters. 2008;101(3):285--7.

\hypertarget{ref-Kang2016}{}
23. Kang H, Zhang A, Cai TT, Small DS. Instrumental Variables Estimation
With Some Invalid Instruments and its Application to Mendelian
Randomization. Journal of the American Statistical Association.
2016;111(513):132--44.

\hypertarget{ref-Steiger1980}{}
24. Steiger JH. Tests for comparing elements of a correlation matrix.
Psychological Bulletin. 1980;87(2):245--51.

\hypertarget{ref-Corbin2016}{}
25. Corbin LJ, Richmond RC, Wade KH, Burgess S, Bowden J, Smith GD, et
al. BMI as a Modifiable Risk Factor for Type 2 Diabetes: Refining and
Understanding Causal Estimates Using Mendelian Randomization. Diabetes
{[}Internet{]}. 2016 Oct;65(10):3002--7. Available from:
\href{http://www.ncbi.nlm.nih.gov/pubmed/27402723\%20http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC5279886}{http://www.ncbi.nlm.nih.gov/pubmed/27402723 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC5279886}

\hypertarget{ref-Lee2013c}{}
26. Lee SH, Wray NR. Novel genetic analysis for case-control genome-wide
association studies: quantification of power and genomic prediction
accuracy. PLoS One. 2013;8(8):e71494.

\hypertarget{ref-Jordan1994}{}
27. Jordan MI, Jacobs RA. Hierarchical Mixtures of Experts and the EM
Algorithm. Neural Computation {[}Internet{]}. 1994 Mar;6(2):181--214.
Available from:
\url{http://www.mitpressjournals.org/doi/10.1162/neco.1994.6.2.181}

\hypertarget{ref-brazdil2008metalearning}{}
28. Brazdil P, Carrier CG, Soares C, Vilalta R. Metalearning:
Applications to data mining. Springer Science \& Business Media; 2008.

\hypertarget{ref-smith2009cross}{}
29. Smith-Miles KA. Cross-disciplinary perspectives on meta-learning for
algorithm selection. ACM Computing Surveys (CSUR). 2009;41(1):6.

\hypertarget{ref-lemke2015metalearning}{}
30. Lemke C, Budka M, Gabrys B. Metalearning: a survey of trends and
technologies. Artificial intelligence review. 2015;44(1):117--30.

\hypertarget{ref-Liaw2002}{}
31. Liaw A, Wiener M. Classification and Regression by randomForest. R
News {[}Internet{]}. 2002;2(3):18--22. Available from:
\url{http://cran.r-project.org/doc/Rnews/}

\hypertarget{ref-Shin2014}{}
32. Shin S-Y, Fauman EB, Petersen A-K, Krumsiek J, Santos R, Huang J, et
al. An atlas of genetic influences on human blood metabolites. Nature
genetics {[}Internet{]}. 2014 Jun;46(6):543--50. Available from:
\url{http://dx.doi.org/10.1038/ng.2982}

\hypertarget{ref-Kettunen2016}{}
33. Kettunen J, Demirkan A, Wurtz P, Draisma HHM, Haller T, Rawal R, et
al. Genome-wide study for circulating metabolites identifies 62 loci and
reveals novel systemic effects of LPA. Nat Commun {[}Internet{]}. 2016
Mar;7. Available from:
\href{http://dx.doi.org/10.1038/ncomms11122\%20http://10.0.4.14/ncomms11122}{http://dx.doi.org/10.1038/ncomms11122 http://10.0.4.14/ncomms11122}

\hypertarget{ref-Sun2017}{}
34. Sun BB, Maranville JC, Peters JE, Stacey D, Staley JR, Blackshaw J,
et al. Consequences Of Natural Perturbations In The Human Plasma
Proteome. bioRxiv {[}Internet{]}. 2017; Available from:
\url{http://biorxiv.org/content/early/2017/05/05/134551}

\hypertarget{ref-Willer2013}{}
35. Willer CJ, Schmidt EM, Sengupta S, Peloso GM, Gustafsson S, Kanoni
S, et al. Discovery and refinement of loci associated with lipid levels.
Nature Genetics {[}Internet{]}. 2013;45(11):1274--83. Available from:
\url{http://www.nature.com/doifinder/10.1038/ng.2797}

\hypertarget{ref-Okbay2016}{}
36. Okbay A, Beauchamp JP, Fontana MA, Lee JJ, Pers TH, Rietveld CA, et
al. Genome-wide association study identifies 74 loci associated with
educational attainment. Nature {[}Internet{]}. 2016
May;533(7604):539--42. Available from:
\href{http://www.ncbi.nlm.nih.gov/pubmed/27225129\%20http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC4883595\%20http://www.nature.com/doifinder/10.1038/nature17671}{http://www.ncbi.nlm.nih.gov/pubmed/27225129 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC4883595 http://www.nature.com/doifinder/10.1038/nature17671}

\hypertarget{ref-Anderson2017}{}
37. Anderson E, Wade KH, Hemani G, Bowden J, Korologou-Linden R, Davey
Smith G, et al. The Causal Effect Of Educational Attainment On
Alzheimer's Disease: A Two-Sample Mendelian Randomization Study. bioRxiv
{[}Internet{]}. 2017; Available from:
\url{http://www.biorxiv.org/content/early/2017/04/17/127993}

\hypertarget{ref-kotthoff2016auto}{}
38. Kotthoff L, Thornton C, Hoos HH, Hutter F, Leyton-Brown K. Auto-WEKA
2.0: Automatic model selection and hyperparameter optimization in WEKA.
Journal of Machine Learning Research. 2016;17:1--5.

\hypertarget{ref-Zhu2016}{}
39. Zhu Z, Zhang F, Hu H, Bakshi A, Robinson MR, Powell JE, et al.
Integration of summary data from GWAS and eQTL studies predicts complex
trait gene targets. Nature Genetics {[}Internet{]}. 2016
Mar;48(5):481--7. Available from:
\url{http://www.nature.com/doifinder/10.1038/ng.3538}

\hypertarget{ref-Richardson2017}{}
40. Richardson TG, Zheng J, Davey Smith G, Timpson NJ, Gaunt TR, Relton
CL, et al. Causal epigenome-wide association study identifies CpG sites
that influence cardiovascular disease risk. bioRxiv {[}Internet{]}.
2017; Available from:
\url{http://biorxiv.org/content/early/2017/04/29/132019}

\hypertarget{ref-Wang2017}{}
41. Wang L, Michoel T. Efficient And Accurate Causal Inference With
Hidden Confounders From Genome-Transcriptome Variation Data. bioRxiv
{[}Internet{]}. 2017; Available from:
\url{http://www.biorxiv.org/content/early/2017/04/19/128496}

\hypertarget{ref-Noyce2017}{}
42. Noyce AJ, Kia DA, Hemani G, Nicolas A, Price TR, De Pablo-Fernandez
E, et al. Estimating the causal influence of body mass index on risk of
Parkinson disease: A Mendelian randomisation study. Brayne C, editor.
PLOS Medicine {[}Internet{]}. 2017 Jun;14(6):e1002314. Available from:
\url{http://dx.plos.org/10.1371/journal.pmed.1002314}

\hypertarget{ref-Burgess2014a}{}
43. Burgess S, Freitag DF, Khan H, Gorman DN, Thompson SG. Using
multivariable Mendelian randomization to disentangle the causal effects
of lipid fractions. PloS one {[}Internet{]}. 2014 Jan;9(10):e108891.
Available from:
\url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0108891}

\hypertarget{ref-Staley2016}{}
44. Staley JR, Blackshaw J, Kamat MA, Ellis S, Surendran P, Sun BB, et
al. PhenoScanner: a database of human genotype--phenotype associations.
Bioinformatics {[}Internet{]}. 2016 Oct;32(20):3207--9. Available from:
\href{http://www.ncbi.nlm.nih.gov/pubmed/27318201\%20http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC5048068\%20https://academic.oup.com/bioinformatics/article-lookup/doi/10.1093/bioinformatics/btw373}{http://www.ncbi.nlm.nih.gov/pubmed/27318201 http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=PMC5048068 https://academic.oup.com/bioinformatics/article-lookup/doi/10.1093/bioinformatics/btw373}

\end{document}