### 2019-02-24 analysis/broad_characteristics.rmd
updated

---
title: Broad characteristics
author: Gibran Hemani
date: 23/02/2019
---

```{r}
library(mrever)
connect()
library(dplyr)
library(data.table)
library(knitr)
library(ggplot2)

traits <- get_traits()
o <- list()
for(i in 1:8)
{
	message(i)

	a <- fread(paste0("zcat ../resources/neo4j_stage/mr", i, ".csv.gz"), he=i == 1, sep=",")
	if(i != 1) names(a) <- names(o[[1]])[-length(names(o[[1]]))]
	names(a) <- gsub(":FLOAT", "", names(a))
	names(a) <- gsub(":INT", "", names(a))
	names(a)[1:2] <- c("exposure", "outcome")
	a$id <- paste(a$exposure, a$outcome)
	a$exposure <- as.character(a$exposure)
	a$outcome <- as.character(a$outcome)
	k <- subset(a, method == "FE IVW" & selection == "Tophits" & nsnp > 5) %>% .$id
	a <- subset(a, paste(exposure, outcome) %in% k)
	o[[i]] <- a
}
o <- bind_rows(o)

moe1 <- subset(o, method == "Steiger null") %>% subset(!duplicated(id))
moe2 <- filter(o, ! id %in% moe1$id) %>% group_by(id) %>% arrange(desc(moescore)) %>% slice(1)

moe <- bind_rows(moe1, moe2)
table(moe$method, moe$selection)
table(moe$method)

moe$z <- abs(moe$b / moe$se)
moe$z[moe$z > 10] <- 10
table(z > 100)

table(a$method, a$selection)

ggplot(moe, aes(x=z)) +
geom_histogram(aes(fill=selection), position="dodge") +
facet_grid(method ~ .) +
theme(strip.text = element_text(angle=0))

```

## 1. Reverse direction results

```{r}

b <- subset(a, id %in% subset(a, method == "Steiger null")$id)
bsig <- subset(b, method == "RE IVW" & selection == "HF" & p.adjust(pval, "fdr") < 0.05)
bsig <- merge(bsig, subset(traits, select=c(bgcidId, trait)), by.x="exposure", by.y="bgcidId")
bsig <- merge(bsig, subset(traits, select=c(bgcidId, trait)), by.x="outcome", by.y="bgcidId")
subset(bsig, select=c(trait.x, trait.y)) %>% kable
```

## 2.

## 1. Compare causal effect estimates across ivw, ivw HF.

What are notable discrepancies?

2. Compare