### 2019-02-27 analysis/ivr2.rmd
updated

---
title: Using the graph
---

1. Trait info is problematic
2. Method is not really applicable when fewer than say 5 SNPs

1. Find instruments for x
2. See if x associates with any other X
3. Get all instruments for X and make an "exposure set"
4. For x-y estimate remove y from exposure set
5.

1. Get list of independent traits
2.

Overcoming reverse causal instruments
- how many reverse causal instruments are there??

Predicting the parameter space improves over single methods

Application to summary data reveals extensive connectedness throughout the phenome

Horizontal pleiotropy increases as more instruments are discovered

```{r}
# library(mrever)
library(igraph)
connect()

mrever_to_mr <- function(id1, id2)
{
	a <- TwoSampleMR::format_data(
		get_instruments(id1),
		snp_col="variantId",
		effect_allele_col="alt",
		other_allele_col="ref",
		id_col="bgcidId",
		units_col="unit"
	)
	b <- TwoSampleMR::format_data(
		get_genassoc(a$SNP, id2),
		type="outcome",
		snp_col="variantId",
		effect_allele_col="alt",
		other_allele_col="ref",
		id_col="bgcidId",
		units_col="unit"
	)
	return(TwoSampleMR::harmonise_data(a,b))
}

fill_in_sd <- function(traits)
{

	estimate_trait_sd <- function(b, se, n, p)
	{
		z <- b / se
		standardised_bhat <- sqrt((z^2/(z^2+n-2)) / (2 * p * (1-p))) * sign(z)
		estimated_sd <- b / standardised_bhat
		return(median(estimated_sd, na.rm=T))
	}
	# table(is.na(traits$sd), traits$unit) %>%t
	traits$unit[is.na(traits$unit)] <- "log odds"

	# For the time being just treat logor as continuous
	traits$SD <- traits$sd
	# table(is.na(traits$SD), traits$unit) %>% t
	traits$unit[traits$unit == "NA" & traits$ncase != 0] <- "risk difference"
	traits$unit[traits$unit == "NA" & traits$ncase == 0] <- "SD"
	traits$SD[grepl("SD", traits$unit)] <- 1

	ind <- traits$unit == "risk difference" & traits$ncontrol == 0
	traits$ncontrol[ind] <- traits$sample_size[ind] - traits$ncase[ind]

	ind <- traits$unit == 'risk difference'
	# traits$SD[ind] <- (traits$ncase[ind] / traits$sample_size[ind]) * (traits$ncontrol[ind] / traits$sample_size[ind])
	traits$SD[ind] <- NA

	rsidlist <- get_instruments(2)$variantId
	todo <- subset(traits, is.na(SD))$bgcidId
	for(i in 1:length(todo))
	{
		message(i)
		inst <- get_instruments(todo[i])
		if(is.null(inst)) inst <- get_genassoc(rsidlist, todo[i])
		miseaf <- inst$variantId[is.na(inst$eaf)]
		if(length(miseaf) > 0)
		{
			eaf <- get_genassoc(miseaf, "UKB-a:100") %>% dplyr::select(variantId, eaf) %>% filter(variantId %in% inst$variantId)
			inst$eaf[match(eaf$variantId, inst$variantId)] <- as.numeric(eaf$eaf)
			inst$eaf <- as.numeric(inst$eaf)
		}
		s <- estimate_trait_sd(inst$beta, inst$se, inst$samplesize, inst$eaf)
		message(s)
		traits$SD[traits$bgcidId == todo[i]] <- s
	}	
	traits$sd <- traits$SD
	return(traits)

}
```

```{r}
# l <- list()
# for(i in 1:nrow(traits))
# {
# 	inst <- get_instruments(traits$bgcidId[i])
# 	if(is.null(inst))
# 		next
# 	cands <- genassocscan(inst$variantId, 0.05 / length(traits) / nrow(inst))

# }

traits <- get_traits() %>% fill_in_sd(.)
traitst <- subset(traits, unit != "mm3")

# Do phewas of everything
l <- lapply(1:nrow(traits), function(i) phewas(traits$bgcidId[i]))
l[sapply(l, is.null)] <- NULL

# Remove traits that have fewer than 80% of results
l[sapply(l, function(x) nrow(x) < length(l) * 0.8)] <- NULL

phe <- bind_rows(l)
save(phe, file="phewas.rdata")

idlist <- unique(phe$exposure)
phe <- subset(phe, outcome %in% exposure)
table(phe$method)
phe$se[is.na(phe$se)] <- 1
phe$se <- abs(phe$se)
phe$z <- pmin(abs(phe$b / phe$se), 10)

tr <- bind_cols(tibble(bgcidId=traits$bgcidId), subset(traits, select=-c(bgcidId))) %>% filter(bgcidId %in% c(l$exposure))

gr <- graph_from_data_frame(subset(l, z > qnorm(1e-16, low=F), select=c(exposure, outcome)), directed=TRUE, vertices = tr)
E(gr)$weight <- subset(l, z > qnorm(1e-16, low=F))$z

hist(l$z)
```

estimate_ivr2 <- function(l, traits)
{
	indx <- match(l$exposure, traits$bgcidId)
	indy <- match(l$outcome, traits$bgcidId)
	l$ivr2 <- l$b^2 * traits$sd[indx]^2 / traits$sd[indy]^2
	return(l)
}

o <- estimate_ivr2(phe, traits)
o <- subset(o, pval < 1e-6)
subset(o, ivr2 > 10) %>% arrange(desc(ivr2)) %>% head

hist(o$ivr2)

o <- subset(o, exposure %in% traitst$bgcidId & outcome %in% traitst$bgcidId)

o %>% arrange(desc(ivr2)) %>% head
hist($ivr2)

wc <- cluster_walktrap(gr, steps=1, weights = E(gr)$weight)
com <- communities(wc)

hs <- hub_score(gr)$vector

plot(gr)

l <- subset(l, )

com <- communities(gr)